`intRo`

### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
working with rmarkdown instead

<style>
.reveal section del,
.reveal h1 del,
.reveal h3 del {
  color: #1abc9c;
}

.reveal, .reveal h3 {
  color: #2c3e50;
}

.reveal section code {
  font-size: 24px;
}

.reveal table {
  font-size: 14px;
  border-collapse: collapse;
  float: left;
}

/*** section background ***/
.section .reveal .state-background {
   background: #2c3e50;
}

.section .reveal a {
  color: #ecf0f1;
}

/*** navigation color ***/
.reveal .controls div.navigate-left,
.reveal .controls div.navigate-left.enabled {
  border-right-color: #1abc9c;
}

.reveal .controls div.navigate-right,
.reveal .controls div.navigate-right.enabled {
  border-left-color: #1abc9c;
}

.reveal .controls div.navigate-up,
.reveal .controls div.navigate-up.enabled {
  border-bottom-color: #1abc9c;
}

.reveal .controls div.navigate-down,
.reveal .controls div.navigate-down.enabled {
  border-top-color: #1abc9c;
}

.reveal .controls div.navigate-left.enabled:hover {
  border-right-color: #148f77;
}

.reveal .controls div.navigate-right.enabled:hover {
  border-left-color: #148f77;
}

.reveal .controls div.navigate-up.enabled:hover {
  border-bottom-color: #148f77;
}

.reveal .controls div.navigate-down.enabled:hover {
  border-top-color: #148f77;
}

.reveal .progress span {
  background: #1abc9c;
}

.reveal blockquote {
  background: #1abc9c;
  color: white;
}
</style>

---
title: "intRo"
author: "Andee Kaplan & Eric Hare"
date: "January 12, 2015"
output: ioslides_presentation
font-family: Helvetica
---

Statistical Analysis Software for Teaching

## Introduction
type: section

## Do we really need another statistical software package?
Short answer: ~~yes~~

@@

## What is intRo?
 - A modern ~~web-based~~ application for performing basic data analysis and statistical routines
 - Built using `R` and `Shiny`
 - Powerful and extensible modular structure
 - Simple enough for the novice statistician
 - Meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education

@@

<img src=images/user_experience.png height="300"></img>

@@

~~Ulterior motive~~: get students excited about programming
 - By navigating about the user interface of intRo, students are creating a fully-executable `R` script that they can download and run locally
 - Viewing their script change real-time within the application

## Extensible
 - User interaction with intRo is split into bitesize chunks that we call *modules*
 - Each module is a self contained set of `R` code that is dynamically added to the application at run time
 - `intRo` can be easily extended by the addition of modules within the frame-work underlying the application
 - Allows `intRo` to be tailored to the needs of a particular course

## First look
type: section

http://intro-stats.com

## Classroom
type: section

## intRo in the wild
 - `intRo` is currently being used in STAT 201
  - Homework assignments and labs will use `intRo` to emphasize class concepts
 - `intRo` is a supplement, not the focus of the class

## Homework 1
 - *Data*: OECD PISA Results in Focus report to assess the workforce readiness of 15-year old students
 - *Goal*: Explore graphically and numerically how the students perceive academics and math
 - `intRo` *Tasks*:
   1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
  1. Make a barchart of the variable Enjoy.Maths.
  1. Give the conditional distribution of ``Strongly agree'' given the Country.
  2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths

## Our turn
type: section

## Design decisions
type: section

@@

<img src="images/ui_annotate.png" height="500"></img>

## Consistent UI (Cont'd)
 - Large, easy to click icons in the page header to help students find
exactly what they need easily
 - Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code
 - Documentation website that is consistent with
the interface of the application; covers all default modules and makes learning to use intRo even more painless

## Modularity
![app creation](images/app_creation_modules.png)

@@

## mylist is a list containing the different ui module code
## Create the UI
shinyUI(
    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
               tabPanel(title="", icon=icon("home"),
                        fluidRow(
                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                        )
               ), ...
    ))
```

@@

## Challenges
type: section

@@

* Storing the user's results
* Presenting only relevant options to the user
* Producing clean and executable R code

## Storing results
* Users must be able to print results, but only the results they want
* Reactivity makes this somewhat challenging

~~Solution~~: Store Buttons

![](images/store.png)

## Storing results (Cont'd)
* Each module contains, at the bottom, a Store button
* When Store is pressed, code is stored in code panel
    * Additional challenge: All options selected must also be stored!
* When user goes to print results, only this particular code is executed

## Presenting choices
* intRo is flexible, but must also be simple for the introductory student
* Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.

~~Example~~:

@@

## Presenting choices (Cont'd)
<img src="images/graphical_table.png">

@@

* Executable on the server
* Executable on the user's machine
* Clean

@@

## Future work
type: section

## What's next?
* Module creation package
* Server load
* Use in other classes

## Module creation package
*Modularity* is a key feature of intRo, but module creation is currently:
* Undocumented
* Entirely manual
* Unnecessarily lengthy

~~Idea~~: R package (with associated Shiny app?) to automate creation of intRo modules

## Server load
Limited testing has been done to assess how intRo handles heavy user load
* Dozens of students at accessing app at once
* Simultaneous larger computations like printing results

More server resources may need to be devoted to handle this, particularly if intRo is more widely adopted

## More classes
* We would like to see intRo used in more classes
* Module creation package will allow it to suit more curricula
* We welcome collaborators interested in extending intRo to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
    * Additional modules
    * Improvements and bug fixes

@@

## Any questions?
type: section

Thank you!

>>>>

---
title: "`intRo`"
subtitle: "Statistical Analysis Software for Teaching"
author: "Andee Kaplan & Eric Hare"
date: "January 12, 2015"
output: ioslides_presentation
font-family: Helvetica
ioslides_presentation:
    logo: images/icon.png
    widescreen: true
---

# Introduction

## Do we really need another statistical software package?
Short answer: **yes**

@@

## What is `intRo`?
 - A modern **web-based** application for performing basic data analysis and statistical routines
 - Built using `R` and `Shiny`
 - Powerful and extensible modular structure
 - Simple enough for the novice statistician
 - Meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education

@@

<img src=images/user_experience.png height="250"></img>

@@

**Ulterior motive**: get students excited about programming
 - By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally
 - Viewing their script change real-time within the application

## Extensible
 - User interaction with `intRo` is split into bitesize chunks that we call *modules*
 - Each module is a self contained set of `R` code that is dynamically added to the application at run time
 - `intRo` can be easily extended by the addition of modules within the frame-work underlying the application
 - Allows `intRo` to be tailored to the needs of a particular course

# First look | http://intro-stats.com

# Classroom

## `intRo` in the wild
 - `intRo` is currently being used in STAT 201
  - Homework assignments and labs will use `intRo` to emphasize class concepts
 - `intRo` is a supplement, not the focus of the class

## Homework 1
 - *Data*: OECD PISA Results in Focus report to assess the workforce readiness of 15-year old students
 - *Goal*: Explore graphically and numerically how the students perceive academics and math
 - `intRo` *Tasks*:
  1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
  1. Make a barchart of the variable Enjoy.Maths.
  1. Give the conditional distribution of "Strongly agree'' given the Country.
  2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths

# Our turn

# Design decisions

@@

<img src="images/ui_annotate.png" height="400"></img>

## Consistent UI (Cont'd)
 - Large, easy to click icons in the page header to help students find
exactly what they need easily
 - Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code
 - Documentation website that is consistent with
the interface of the application; covers all default modules and makes learning to use `intRo` even more painless

## Modularity
<img src="images/app_creation_modules.png" height="500"></img>

@@

## mylist is a list containing the different ui module code
## Create the UI
shinyUI(
    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
               tabPanel(title="", icon=icon("home"),
                        fluidRow(
                        ### <b>
                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                        ### </b>
                        )
               ), ...
    ))
```

@@

# Challenges

@@

 - Storing the user's results
 - Presenting only relevant options to the user
 - Producing clean and executable R code

## Storing results
 - Users must be able to print results, but only the results they want
 - Reactivity makes this somewhat challenging

**Solution**: Store Buttons

<img src="images/store.png" height="300"></img>

## Storing results (Cont'd)
 - Each module contains, at the bottom, a Store button
 - When Store is pressed, code is stored in code panel
     - Additional challenge: All options selected must also be stored!
 - When user goes to print results, only this particular code is executed

## Presenting choices
 - `intRo` is flexible, but must also be simple for the introductory student
 - Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.

**Example**:

@@

## Presenting choices (Cont'd)
<img src="images/graphical_table.png" height="275">

@@

 - Executable on the server
 - Executable on the user's machine
 - Clean

@@

# Future work

## What's next?
 - Module creation package
 - Server load
 - Use in other classes

## Module creation package
*Modularity* is a key feature of `intRo`, but module creation is currently:
 - Undocumented
 - Entirely manual
 - Unnecessarily lengthy

**Idea**: R package (with associated Shiny app?) to automate creation of `intRo` modules

## Server load
Limited testing has been done to assess how `intRo` handles heavy user load
 - Dozens of students at accessing app at once
 - Simultaneous larger computations like printing results

More server resources may need to be devoted to handle this, particularly if `intRo` is more widely adopted

## More classes
 - We would like to see `intRo` used in more classes
 - Module creation package will allow it to suit more curricula
 - We welcome collaborators interested in extending `intRo` to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
     - Additional modules
     - Improvements and bug fixes

@@

# Any questions? | Thank you!
